### 2011-11-22 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: fixing singletons and merging Kesy.

We shall use the representation introduced in \cite{fidax}, where an XML file is represented by a labeled tree consisting of nodes for elements, attributes and simple text data. Parent nodes are connected to child nodes with edges. This tree shall be called an \textit{XML tree}. For a given node $v$ of an XML tree we define $label(v)$ (name of the node in in the document, only for elements and attributes), $id(v)$ (unique identifier across the document) and $value(v)$ (text content, only for attributes and simple text data) in the same way as the cited article does.

@@

We can use projection to retrieve all the unique \textit{ids} of either elements or attributes from the relation, with notation $\pi_E(M_{x}^{y})$ and $\pi_A(M_{x}^{y})$.

@@

\paragraph{Example}
Returning to our example, the following are all the possible ID sets: $\{ M_{y}^{a} \}, \{ M_{y}^{b} \}, \{ M_{y}^{b}, M_{z}^{a} \}, \{ M_{y}^{c} \}, \{ M_{y}^{c}, M_{z}^{a} \}$. Note that once we select an AM of type $y$ we can never add any other with the same type. Note also that $\{ M_{y}^{c}, M_{z}^{a} \}$ is not an ID set, because $\iota(M_{y}^{c}) \cap \iota(M_{z}^{a}) \neq \emptyset$.

@@

\begin{define}[Support]
\textit{Support} of an attribute mapping $m$ is defined as follows:
\[\phi(m) = \frac{|m|}{\sum_{p \in M}|p|}.\]
\end{define}

@@

\begin{define}[Coverage]
\textit{Coverage} of an attribute mapping $m$ is defined as follows:
\[\chi(m) = \left( \sum_{p \in M, p \neq m} |\iota(m) \cap \iota(p)| \right) / \sum_{p \in M} |\iota(p)|.\]
\end{define}

@@

\begin{define}[Maximum weighted independent set]
Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an independent set $I_{max}$ is the \textit{maximum weighted independent set}, iff the following is satisfied:
\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v \in I_{max}} w(v).\]
\end{define}

@@

Where $\mathbf{x}$ is the vector of variables (to be found by the optimization), $\mathbf{b}$ is the vector and $A$ its accompanying matrix of constraints and $\mathbf{c}$ is the vector of coefficients for the objective function. $\mathbf{x}$ and $\mathbf{c}$ have length $n$, $\mathbf{b}$ has length $m$ and $A$ has dimensions $m \times n$. Furthermore, \[\max_{x} f(x)\] means maximize the value of $f(x)$ by changing the value of $x$.\\

>>>>

We shall use the representation introduced in \cite{fidax}, where an XML file is~represented by a labeled tree consisting of nodes for elements, attributes and simple text data. Parent nodes are connected to child nodes with edges. This tree shall be called an \textit{XML tree}. For a given node $v$ of an XML tree we define $label(v)$ (name of the node in in the document, only for elements and attributes), $id(v)$ (unique identifier across the document) and $value(v)$ (text content, only for attributes and simple text data) in the same way as the cited article does.

@@

We can use projection to retrieve all the unique \textit{ids} of either elements or~attributes from the relation, with notation $\pi_E(M_{x}^{y})$ and $\pi_A(M_{x}^{y})$.

@@

\paragraph{Example}
Returning to our example, the following are all the possible ID sets: $\{ M_{y}^{a} \}, \{ M_{y}^{b} \}, \{ M_{y}^{b}, M_{z}^{a} \}, \{ M_{y}^{c} \}, \{ M_{y}^{c}, M_{z}^{a} \}$. Note that once we select an AM of~type $y$ we can never add any other with the same type. Note also that $\{ M_{y}^{c}, M_{z}^{a} \}$ is~not an ID set, because $\iota(M_{y}^{c}) \cap \iota(M_{z}^{a}) \neq \emptyset$.

@@

\begin{define}[Support]
\textit{Support} of an attribute mapping $m$ is defined as~follows:
\[\phi(m) = \frac{|m|}{\sum_{p \in M}|p|}.\]
\end{define}

@@

\begin{define}[Coverage]
\textit{Coverage} of an attribute mapping $m$ is defined as~follows:
\[\chi(m) = \left( \sum_{p \in M, p \neq m} |\iota(m) \cap \iota(p)| \right) / \sum_{p \in M} |\iota(p)|.\]
\end{define}

@@

\begin{define}[Maximum weighted independent set]
Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an~independent set $I_{max}$ is~the \textit{maximum weighted independent set}, iff the following is satisfied:
\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v \in I_{max}} w(v).\]
\end{define}

@@

Where $\mathbf{x}$ is~the vector of variables (to be found by the optimization), $\mathbf{b}$ is~the vector and $A$ its accompanying matrix of constraints and $\mathbf{c}$ is the vector of coefficients for the objective function. $\mathbf{x}$ and $\mathbf{c}$ have length $n$, $\mathbf{b}$ has length $m$ and $A$ has dimensions $m \times n$. Furthermore, \[\max_{x} f(x)\] means maximize the value of $f(x)$ by changing the value of $x$.\\

### 2011-11-22 doc/theses/vitasek/chapter-2-research.tex
Thesis: fixing singletons and merging Kesy.

\begin{quote}
The idea behind is an observation that a projection of entities corresponds to a key if each counted aggregation for a projection is equal to 1. Thus, this method searches for all possible projections of a dataset while computing aggregations on the projected part of the set of entities.
\end{quote}

@@

Maximum weigthed IS is a well researched topic with a lot of known direct or approximation algorithms, see e.g. \cite{JM1986425} or \cite{Fomin:2009:MCA:1552285.1552286}. According to for example \cite{Paschos:1997:SAO:254180.254190}, the best known approximation algorithm for weighted IS to-date achieves an approximation ratio of $3(\Delta + 2)$, where $\Delta$ is the maximum degree of a vertex in the IS graph. This article lists several algorithms similar to those we introduce in the following chapters.\\

>>>>

\begin{quote}
The idea behind is~an~observation that a~projection of~entities corresponds to~a~key if~each counted aggregation for a~projection is~equal to 1. Thus, this method searches for all possible projections of a~data\-set while computing aggregations on~the projected part of~the set of~entities.
\end{quote}

@@

Maximum weigthed IS is a well researched topic with a lot of~known direct or~approximation algorithms, see e.g. \cite{JM1986425} or \cite{Fomin:2009:MCA:1552285.1552286}. According to \cite{Paschos:1997:SAO:254180.254190}, the best known approximation algorithm for weighted IS to-date achieves an~approximation ratio of $3(\Delta + 2)$, where $\Delta$ is the maximum degree of a vertex in~the IS graph. This article lists several algorithms similar to those we introduce in~the following chapters.\\

### 2011-11-22 doc/theses/vitasek/chapter-3-mip.tex
Thesis: fixing page numbers and a figure.
Thesis: fixing singletons and merging Kesy.

GLPK is a multi-platform, multi-purpose solver well suited for this task. It uses the Simplex method to solve LP problems and \textit{Branch \& Bound} for MIP.

\paragraph{\textit{Branch \& Bound}} This is an optimization method where the search space is systematically divided into smaller sub-spaces. This is the \textit{branching} component, and a so-called \textit{search tree} is built this way. Then, the sub-problems are recursively solved and whole branches of the search tree are discarded when it becomes obvious that the solution does not lie there. This is the \textit{bounding} component. See \cite{land60a} for detailed description.\\

@@

\begin{figure}
  \centering
	\includegraphics[width=.25\textwidth]{images/is-representation}
  \caption{IS Representation Graph}
  \label{image-mip-is-graph}
\end{figure}

@@

\begin{define}[Solution Space, Solution Quality]
	Solution space in general is the set of all permissible solutions (not violating any constraints). In the specific case of a MIP formulation it is the set of all $\mathbf{x}$ subject to $A\mathbf{x} \leqslant \mathbf{b}$. Every solution in the solution space has its \textit{quality}, in case of MIP for solution $\mathbf{x}$ it is the value of the objective function in $\mathbf{x}$.
\end{define}

@@

A quick reminder of what we are trying to solve using our metaheuristic: given a list of AMs with weights, find a non-conflicting subset maximizing the sum of weights in the subset. We will now describe its structure, please refer again to Figure \ref{image-metaheuristic}.

@@

The algorithm works in two steps. First, all candidate AMs are grouped according to their types, and for each type the AM with the highest weight is selected. Second, all these AMs (now called $C'$) are traversed in order of their decreasing size. For each AM $m$, a set $S$ of all conflicting AMs from $C'$ is found and weights of both $m$ and $S$ are calculated. Then the weights are compared and either $m$ or $S$ is removed from $C'$.

@@

One of the most natural heuristics when dealing with the ID set problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution. This is obviously a greedy heuristic.

@@

\heu{Fuzzy} is an improvement over the \heu{Random} CH: the next AM to be added is selected based on \textit{weighted} instead of uniform random. The weight used here is the usual weight of an AM as defined in Section \ref{section-definitions-weight}. Because of the randomness involved in the choice, we can again easily create a pool of solutions this way.

>>>>

GLPK is a multi-platform, multi-purpose solver well suited for this task. It~uses the Simplex method to solve LP problems and \textit{Branch \& Bound} for MIP.

\paragraph{\textit{Branch \& Bound}} This is an optimization method where the search space is~systematically divided into smaller sub-spaces. This is the \textit{branching} component, and a so-called \textit{search tree} is built this way. Then, the sub-problems are recursively solved and whole branches of the search tree are discarded when it becomes obvious that the solution does not lie there. This is the \textit{bounding} component. See \cite{land60a} for detailed description.\\

@@

\begin{figure}
  \caption{IS Representation Graph}
  \label{image-mip-is-graph}
  \centering
	 \includegraphics[width=.25\textwidth]{images/is-representation}
\end{figure}

@@

\begin{define}[Solution Space, Solution Quality]
	Solution space in general is~the~set of all permissible solutions (not violating any constraints). In the specific case of a MIP formulation it is the set of all $\mathbf{x}$ subject to $A\mathbf{x} \leqslant \mathbf{b}$. Every solution in the solution space has its \textit{quality}, in case of MIP for solution $\mathbf{x}$ it~is~the value of the objective function in $\mathbf{x}$.
\end{define}

@@

A quick reminder of what we are trying to solve using our metaheuristic: given a list of AMs with weights, find a non-conflicting subset maximizing the sum of weights in the subset. This sum will henceforth be considered the quality of the solution (subset). We will now describe its structure, please refer again to Figure \ref{image-metaheuristic}.

@@

The algorithm works in two steps. First, all candidate AMs are grouped according to their types, and for each type the AM with the highest weight is~selected. Second, all these AMs (now called $C'$) are traversed in order of their decreasing size. For each AM $m$, a set $S$ of all conflicting AMs from $C'$ is found and weights of both $m$ and $S$ are calculated. Then the weights are compared and either $m$ or $S$ is removed from $C'$.

@@

One of the most natural heuristics when dealing with the ID set problem can be described as follows: select from candidate AMs at random, if possible (addition would not violate the ID set condition) add them to the solution. This is~obviously a greedy heuristic.

@@

\heu{Fuzzy} is an improvement over the \heu{Random} CH: the next AM to be added is~selected based on \textit{weighted} instead of uniform random. The weight used here is~the usual weight of an AM as defined in Section \ref{section-definitions-weight}. Because of the randomness involved in the choice, we can again easily create a pool of solutions this way.

### 2011-11-22 doc/theses/vitasek/chapter-4-1-experiments-setup.tex
Thesis: fixing singletons and merging Kesy.

In case of realistic data we want to see the performance in cases taken from the real world. The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as simple text nodes instead of attributes. We will try to convert some of these values to attributes (e.g. using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see Section \ref{section-realistic-converted} for details).
And finally, we create completely artificial data to create inputs that will put our heuristics in stress. This is because the realistic data often prove to be too simple to solve - the list of candidate AMs is usually too short to be hard to be solved to optimality.

\begin{define}[Data set]
	\label{define-data-set}
	One or more XML files sharing the same schema (even if only an implicit schema) shall be referred to as a \textit{data set}. In the scope of this work this will always mean a \textit{single} XML file. However, this definition of a data \textit{set} covers also the extension to more XML files as described in \cite{fidax}.
\end{define}

@@

The former will be achieved with the help of the GraphViz tool (\cite{graphviz}), where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of their images (and thus cannot be in the same ID set together). Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.

The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ (found by running the \heu{Glpk} heuristic without a time limit) and the numbers of vertices and edges in aforementioned graphs.

@@

In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was indeed achieved. There is an interesting observation to~be~made here: the optimum is steadily decreasing with the increasing overall graph size. This intuitively suggests that the maximum quality theoretically achievable is related to the $\frac{|E|}{|V|}$ density, not to the one we fixed. Exploration of this phenomenon is beyond the scope of this thesis.\\

@@

An input data set is a single XML file on the filesystem; however, there is a straightforward extension to multiple files conforming to the same schema. The first step in this process is to use jInfer's module \jmodule{BasicIGG} module (see \cite{basiciggdoc} for details) to obtain a list of rules - an \textit{initial grammar} (IG). \nomenclature{IG}{Initial Grammar}
Please see \cite{archdoc} for detailed specification of IG format.\\

>>>>

In case of realistic data we want to see the performance in cases taken from the real world. The problem with realistic data is that sometimes, interesting values (that might or might not contain IDs) are stored as simple text nodes instead of attributes. We will try to convert some of these values to attributes (e.g. using a smart XSL transformation), let our heuristics find the ID sets, and then translate them back to XML keys (see Section \ref{section-realistic-converted} for details).
And finally, we create completely artificial data to create inputs that will put our heuristics in~stress. This is because the realistic data often prove to be too simple to solve - the list of candidate AMs is usually too short to be hard to be solved to optimality.

\begin{define}[Data set]
	\label{define-data-set}
	One or more XML files sharing the same schema (even if only an implicit schema) shall be referred to as a \textit{data set}. In the scope of this work this will always mean a \textit{single} XML file. However, this definition of~a~data \textit{set} covers also the extension to more XML files as described in \cite{fidax}.
\end{define}

@@

The former will be achieved with the help of the GraphViz tool (\cite{graphviz}), where we will draw the graphs so that all the vertices represent the \textit{candidate AMs}, and the edges represent pairs of AMs that have nonempty intersection of~their images (and thus cannot be in the same ID set together). Thus solving the maximal weighted IS on these graphs will be equivalent to solving our problem of optimal ID set.

The latter will come in form of tables containing information regarding the data sets, such their size, known optimum for $\alpha = \beta = 1$ (found by running the \heu{Glpk} heuristic without a time limit) and the numbers of vertices and edges in~aforementioned graphs.

@@

In the second table we aimed for the $\frac{|E|}{E_{max}}$ density of $0.1 = 10\%$, and we can see that this was indeed achieved. There is an interesting observation to~be~made here: the optimum is steadily decreasing with the increasing overall graph size. This intuitively suggests that the maximum quality theoretically achievable is~related to the $\frac{|E|}{|V|}$ density, not to the one we fixed. Exploration of this phenomenon is beyond the scope of this thesis.\\

@@

An input data set is a single XML file on the filesystem; however, there is~a~straightforward extension to multiple files conforming to the same schema. The first step in this process is to use jInfer's module \jmodule{BasicIGG} module (see \cite{basiciggdoc} for details) to obtain a list of rules - an \textit{initial grammar} (IG). \nomenclature{IG}{Initial Grammar}
Please see \cite{archdoc} for detailed specification of IG format.\\
